---
title: "hw9"
author: "XianjieQue"
date: "2025-04-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

### load packages

```{r load packages}
library(ggplot2)
library(ISLR2)
library(boot)
library(splines)
library(leaps)
library(glmnet)
# install.packages("gam")
library(gam)
```

Section 7.9

### Conceptual: 1

What I need to do is to show that a function of the form\
$$
f(x) = \beta_0 + \beta_1 x + \beta_2 x^2 + \beta_3 x^3 + \beta_4 (x - \xi)^3_+
$$\
is a cubic regression spline with one knot at $\xi$.

#### (a)

Define $f_1(x)$ for $x \leq \xi$

When $x \leq \xi$, the truncated cubic term disappears:\
$$
(x - \xi)^3_+ = 0
$$\
So the function becomes: $$
f_1(x) = \beta_0 + \beta_1 x + \beta_2 x^2 + \beta_3 x^3
$$\
Thus, we can write it as a cubic polynomial:\
$$
f_1(x) = a_1 + b_1 x + c_1 x^2 + d_1 x^3
$$\
where:\
$$a_1 = \beta_0 \\ b_1 = \beta_1 \\ c_1 = \beta_2 \\ d_1 = \beta_3$$

#### (b)

Define $f_2(x)$ for $x > \xi$

For $x > \xi$, we expand the truncated cubic term: $$
(x - \xi)^3 = x^3 - 3\xi x^2 + 3\xi^2 x - \xi^3
$$

So: $$
f(x) = \beta_0 + \beta_1 x + \beta_2 x^2 + \beta_3 x^3 + \beta_4(x - \xi)^3\\
= (\beta_0 - \beta_4 \xi^3) + (\beta_1 + 3 \beta_4 \xi^2)x + (\beta_2 - 3 \beta_4 \xi) x^2 + (\beta_3 + \beta_4) x^3
$$

This gives us another cubic polynomial: 
$$
f_2(x) = a_2 + b_2 x + c_2 x^2 + d_2 x^3
$$

where:
$$
a_2 = \beta_0 - \beta_4 \xi^3 \\ b_2 = \beta_1 + 3 \beta_4 \xi^2 \\
c_2 = \beta_2 - 3 \beta_4 \xi \\ d_2 = \beta_3 + \beta_4
$$

#### (c)

Show that $f_1(\xi) = f_2(\xi)$, which means continuity at $\xi$

Compute: 
$$
f_1(\xi) = \beta_0 + \beta_1 \xi + \beta_2 \xi^2 + \beta_3 \xi^3
$$

and: $$
f_2(\xi) = a_2 + b_2 \xi + c_2 \xi^2 + d_2 \xi^3\\
= (\beta_0 - \beta_4 \xi^3) + (\beta_1 + 3 \beta_4 \xi^2) \xi + (\beta_2 - 3 \beta_4 \xi) \xi^2 + (\beta_3 + \beta_4) \xi^3
$$ After simplification, $f_2(\xi) = f_1(\xi)$\
So the function is continuous at $\xi$

#### (d)

Show that $f_1'(\xi) = f_2'(\xi)$, which means first derivative
continuity

Derivative of a cubic: $$
f'(x) = \beta_1 + 2\beta_2 x + 3\beta_3 x^2 + 3\beta_4(x - \xi)^2_+
$$

At $x = \xi$, the last term is zero, so: $$
f_1'(\xi) = f_2'(\xi) = \beta_1 + 2\beta_2 \xi + 3\beta_3 \xi^2
$$ First derivative is continuous at $\xi$.

#### (e)

Show that $f_1''(\xi) = f_2''(\xi)$, which means second derivative
continuity

Second derivative of a cubic: $$
f''(x) = 2\beta_2 + 6\beta_3 x + 6\beta_4(x - \xi)^1_+
$$

At $x = \xi$, the last term is again zero: $$
f_1''(\xi) = f_2''(\xi) = 2\beta_2 + 6\beta_3 \xi
$$

Second derivative is continuous at $\xi$.

All conditions are satisfied for a cubic spline:

1.  Piecewise cubic on either side of $\xi$
2.  Function, first derivative, and second derivative are all continuous
    at $\xi$

$$
{f(x) \text{ is a cubic regression spline with one knot at } \xi}
$$

### Conceptual: 2

#### (a) λ=∞,m=0

We penalize the function $g(x)$ itself. An infinite $\lambda$ means that
this penalty dominates. This forces $\hat{g}$ to be 0 everywhere.

To avoid an infinite penalty, the only solution is $\hat{g}(x) = 0$

#### (b) λ=∞,m=1

We penalize the square of the first derivative $g'(x)$. With
$\lambda = \infty$, the curve must be completely flat to avoid infinite
penalty.

We need $g'(x) = 0$ everywhere. This means $\hat{g}(x)$ must be a
constant function.

Thus, $\hat{g}$ is a constant function: $\hat{g}(x) = c$

#### (c) λ=∞,m=2

We penalize the second derivative $g''(x)$. So the curve must be linear
in order to have zero penalty. Therefore, $\hat{g}$ is a linear
function: $\hat{g}(x) = a x + b$

#### (d) λ=∞,m=3

We penalize the third derivative $g^{(3)}(x)$. To avoid any penalty, the curve will have a consistent rate of change (a quadratic
function or similar). So, $\hat{g}(x)$ can be the form:
$\hat{g}(x) = a x^2 + b x + c$

#### (e) λ=0,m=3

There is no penalty on the third derivative or any smoothness. The
optimization only minimizes the squared errors at the data points.

$\hat{g}$ interpolates all data points exactly, and may be highly
non-smooth. It can be thought of as a high-degree polynomial or a
piecewise spline that satisfies:
$\hat{g}(x_i) = y_i \quad \text{for all } i = 1, \dots, n$

### Applied: 6

#### (a)

CV (cross-validation) selects degree 5, while when testing with ANOVA,
degrees 2 and 3 are highly significant, degree 4 is marginal.

Because CV chooses the degree that gives the lowest average prediction
error on unseen data, while ANOVA tests whether adding higher-degree
terms significantly improves model fit over simpler ones.

```{r 6 (a)}
data(Wage)
sum(is.na(Wage))

# CV
set.seed(9)

cv_error_poly <- sapply(1:5, function(i) {
  fit <- glm(wage ~ poly(age, i), data = Wage)
  cv.glm(Wage, fit, K = 10)$delta[1]
})
best_degree <- which.min(cv_error_poly)
print(paste("Best degree:", best_degree))

plot(1:5, cv_error_poly, pch = 19, type = "b", xlab = "degree of polynomial", ylab = "CV estimate of the prediction error")
grid()

# ANOVA comparison
lm.fit1 <- lm(wage ~ poly(age, 1), data = Wage)
lm.fit2 <- lm(wage ~ poly(age, 2), data = Wage)
lm.fit3 <- lm(wage ~ poly(age, 3), data = Wage)
lm.fit4 <- lm(wage ~ poly(age, 4), data = Wage)
lm.fit5 <- lm(wage ~ poly(age, 5), data = Wage)
anova(lm.fit1, lm.fit2, lm.fit3, lm.fit4, lm.fit5)

# Plot best polynomial fit (degree = 3)
agelims <- range(Wage$age)
age.grid <- seq(from = agelims[1], to = agelims[2])
preds <- predict(lm.fit3, newdata = data.frame(age = age.grid), se = TRUE)
se.bands <- cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)

plot(Wage$age, Wage$wage, xlim = agelims, cex = 0.5, xlab = "Age", ylab = "Wage")
title("Degree-3 Polynomial Fit")
lines(age.grid, preds$fit, lwd = 2, col = "blue")
matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
```

#### (b) Step function with CV

The best number of cuts choosed by CV is 8

```{r 6 (b)}
set.seed(9)

cv_error_step <- rep(NA, 9)
for (i in 2:10) {
  Wage$age.cut <- cut(Wage$age, breaks = i)
  fit <- glm(wage ~ age.cut, data = Wage)
  cv_error_step[i - 1] <- cv.glm(Wage, fit, K = 10)$delta[1]
}
best_cut <- which.min(cv_error_step) + 1
print(paste("Best number of cuts:", best_cut))

# Fit step function with best number of cuts
breaks <- quantile(Wage$age, probs = seq(0, 1, length.out = best_cut + 1))
Wage$age.cut <- cut(Wage$age, breaks = breaks, include.lowest = TRUE)
step.fit <- glm(wage ~ age.cut, data = Wage)

# Plot step function fit
age.grid <- seq(from = min(Wage$age), to = max(Wage$age), length.out = 100)

# Apply same cuts to new data
newdata <- data.frame(age = age.grid)
newdata$age.cut <- cut(newdata$age, breaks = breaks, include.lowest = TRUE)

# Predict using step function
step.pred <- predict(step.fit, newdata = newdata)

# Plot
plot(wage ~ age, data = Wage, xlab = "Age", ylab = "Wage")
title("Step Function Fit")
lines(age.grid, step.pred, col = "red", lwd = 2)
```

### Applied: 9

#### (a)

All terms of the cubic polynomial are significant (p-values are less
than 0.001). This shows that the relationship between dis and nox is not
a simple linear relationship, but has a nonlinear trend.

The nonlinear relationship in the plot shows as: when "dis" increases,
"nox" first decreases rapidly, then stabilizes, and then decreases
slightly.

Multiple R-squared: 0.7148, Adjusted R-squared: 0.7131

The goodness of fit is high, indicating that the polynomial model
adequately explains the data, explaining approximately 71.5% of the
variance.

```{r 9 (a)}
data(Boston)

fit_a <- lm(nox ~ poly(dis, 3), data = Boston)
summary(fit_a)

plot(nox ~ dis, data = Boston, 
     xlab = "dis", ylab = "nox", 
     main = "third degree polynomial fit")

dis_range <- range(Boston$dis)
dis_samples <- seq(from = dis_range[1], to = dis_range[2], length.out = 100)
y_hat <- predict(fit_a, newdata = list(dis = dis_samples))

lines(dis_samples, y_hat, col = "red")
grid()
```

#### (b)

From the output, as the degree increased, the RSS decreased steadily,
indicating improved fit on the training data. However, the marginal
reduction in RSS became negligible after degree 3, suggesting that
increasing complexity beyond that point may lead to overfitting.

```{r 9 (b)}
d_max <- 10

# Training RSS for degrees 1 to 10
training_rss <- rep(NA, d_max)
for (d in 1:d_max) {
    m <- lm(nox ~ poly(dis, d), data = Boston)
    training_rss[d] <- sum((m$residuals)^2)
}

plot(1:d_max, training_rss, type = "b", pch = 19, xlab = "Degree", ylab = "Training RSS")
title("Training RSS vs Polynomial Degree")

# Output the RSS for each degree
rss_df <- data.frame(Degree = 1:10, Training_RSS = training_rss)
print(rss_df)
```

#### (c)

The plot shows that the CV error initially decreases slightly and
reaches its minimum around degree 3. After that, the error begins to
increase, especially at degrees 7, 9, and 10, indicating potential
overfitting.

The optimal degree selected by CV would likely be degree 3. This result aligns with the RSS plot in (b), which showed decreasing returns in training error beyond degree 3.

```{r 9 (c)}
set.seed(9)

d_max <- 10
cv_error <- rep(NA, d_max)

for (d in 1:d_max) {
  glm.fit <- glm(nox ~ poly(dis, d), data = Boston)
  cv_error[d] <- cv.glm(Boston, glm.fit, K = 10)$delta[1]
}

plot(1:d_max, cv_error, type = "b", pch = 19,
     xlab = "Degree", ylab = "CV Error",
     main = "CV Error for Polynomial Degree")

which.min(cv_error)
```

#### (d)

Output:

Using 4 df, all terms are significant.

Multiple R-squared: 0.7164, Adjusted R-squared: 0.7142

The goodness of fit is high, indicating that the polynomial model
adequately explains the data, explaining approximately 71.6% of the
variance.

How to choose the knots?

Knots are chosen based on quantiles of the data.

df = #knots + degree and the default degree is 3 (cubic)

So, bs(dis, df = 4) automatically selects 1 knot over the range of "dis".

I use "attr(bs(Boston\$dis, df = 4), "knots")" to check the knot, the
output "3.20745" means the knot is at "dis = 3.20745"

```{r 9 (d)}
fit_d <- lm(nox ~ bs(dis, df = 4), data = Boston)
summary(fit_d)

attr(bs(Boston$dis, df = 4), "knots")

plot(Boston$dis, Boston$nox, xlab = "dis", ylab = "nox", 
     main = "bs with df=4 fit")

dis_range <- range(Boston$dis)
dis_samples <- seq(from = dis_range[1], to = dis_range[2], length.out = 100)
y_hat <- predict(fit_d, newdata = data.frame(dis = dis_samples))

lines(dis_samples, y_hat, col = "red")
grid()
```

#### (e)

From the result, as the df increased from 3 to 20, RSS decreased from
1.93 to 1.78.

The reduction in RSS was most significant when increasing from 3 to 10
df.

Beyond df = 10, the gain was minimal, suggesting that higher flexibility
may lead to overfitting without substantial improvement.

Therefore, 10 degrees of freedom may offer a good trade-off between bias
and variance in this case.

The same in the plot: The curve with low df (such as df = 3) is too
smooth and cannot capture the nonlinear structure between dis-nox. The
curve with high df (such as df = 15, 20) fits more finely, but may not
perform well on the test set.

```{r 9 (e)}
dof_choices <- c(3, 4, 5, 10, 15, 20)
rss_values <- numeric(length(dof_choices))

plot(Boston$dis, Boston$nox, xlab = "dis", ylab = "nox", 
     main = "Regression Spline Fits")

colors <- rainbow(length(dof_choices))
legend_labels <- c()

for (i in seq_along(dof_choices)) {
  df_val <- dof_choices[i]
  fit <- lm(nox ~ bs(dis, df = df_val), data = Boston)
  
  rss_values[i] <- sum(residuals(fit)^2)
  
  dis.grid <- seq(min(Boston$dis), max(Boston$dis), length.out = 100)
  preds <- predict(fit, newdata = data.frame(dis = dis.grid))
  lines(dis.grid, preds, col = colors[i], lwd = 2)
  
  legend_labels[i] <- paste("df =", df_val)
}

legend("topright", legend = legend_labels, col = colors, lwd = 2)
print(data.frame(dof = dof_choices, RSS = round(rss_values, 2)))
```

#### (f)

I performed cross-validation and the best degrees of freedom based on CV
is 10.

This result is consistent with the RSS trend in (e), where the RSS
plateaued after df = 10. Therefore, df = 10 appears to be an appropriate
complexity level for modeling the relationship between "dis" and "nox".

```{r 9 (f)}
set.seed(9)

dof_choices <- c(3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
cv_errors <- numeric(length(dof_choices))

for (i in seq_along(dof_choices)) {
  df_val <- dof_choices[i]
  glm.fit <- glm(nox ~ bs(dis, df = df_val), data = Boston)
  cv_result <- cv.glm(Boston, glm.fit, K = 10)
  cv_errors[i] <- cv_result$delta[1]
}

plot(dof_choices, cv_errors, type = "b", pch = 19,
     xlab = "Degrees of Freedom", ylab = "10-fold CV Error",
     main = "CV Error vs Spline Degrees of Freedom")

best_df <- dof_choices[which.min(cv_errors)]
points(best_df, min(cv_errors), col = "red", pch = 19, cex = 1.5)
text(best_df, min(cv_errors), labels = paste("Best df:", best_df), pos = 3, col = "red")

print(paste("Best degrees of freedom based on CV:", best_df))
```

### Applied: 10

#### (a)

There are 17 Variables (and intercept) in all.

The model selection criterion I used was BIC.

Best model size choosen by BIC is 10, with coefficients given by: 

"PrivateYes" "Apps" "Accept" "Top10perc" "F.Undergrad" "Room.Board"
"Terminal" "perc.alumni" "Expend" "Grad.Rate"

```{r 10 (a)}
data("College")
sum(is.na(College))

# split 70% as training, 30% as test
set.seed(9)
train <- sample(1:nrow(College), 0.7*nrow(College))

College_train <- College[train,]
College_test <- College[-train,]

x <- College[, !names(College) %in% "Outstate"]
y <- College$Outstate
x_train <- as.matrix(College_train[, !names(College_train) %in% "Outstate"])
y_train <- as.numeric(College_train$Outstate)
x_test <- as.matrix(College_test[, !names(College_test) %in% "Outstate"])
y_test <- as.numeric(College_test$Outstate)

fit_forward <- regsubsets(Outstate ~ ., data = College_train, 
                             nvmax = ncol(College_train) - 1, method = "forward")
fit_summary <- summary(fit_forward)
fit_summary

# Identify best model by BIC
best_model_size <- which.min(fit_summary$bic)
best_model_vars <- names(coef(fit_forward, best_model_size))[-1]  # remove intercept
print(paste("Best model size (by BIC):", best_model_size))
print("Selected predictors:")
print(best_model_vars)


```

#### (b)

Each plot shows the estimated smooth function (s(variable)) of a
predictor on the "Outstate", with 95% CIs (dashed lines).

"PrivateYes" is a categorical variable. There is a large vertical jump
between No and Yes, suggesting that private colleges tend to have
significantly higher out-of-state tuition than public ones.

Variables like "Expend" and "Grad.Rate" display smooth, non-linear
influences.

The GAM captures both linear and non-linear relationships, showing its
flexibly model complex effects without assuming strict linearity.

```{r 10 (b)}
numeric_vars <- names(College)[sapply(College, is.numeric)]

best_model_vars[best_model_vars == "PrivateYes"] <- "Private"

formula_terms <- sapply(best_model_vars, function(var) {
  if (var %in% numeric_vars) {
    paste0("s(", var, ")")
  } else {
    var
  }
})

gam_formula <- as.formula(paste("Outstate ~", paste(formula_terms, collapse = " + ")))

# Fit GAM
fit_gam <- gam(gam_formula, data = College_train)

# Plot the GAM fit
par(mfrow = c(1, 2))
plot(fit_gam, se = TRUE, col = "blue")
```

#### (c)

Test MSE of the GAM: 3906866.28

This means the average squared error between the predicted and actual
Outstate tuition on the test set is about 3.91 million. Since the
tuition values are mostly between 5,000 and 20,000, this MSE suggests
moderate prediction error — but not excellent.

```{r 10 (c)}
pred_gam <- predict(fit_gam, newdata = College_test)

# Compute test MSE
mse_gam <- mean((College_test$Outstate - pred_gam)^2)
print(paste("Test MSE of the GAM:", round(mse_gam, 2)))
```

#### (d)

GAM tests for non-linearity using the ANOVA for Nonparametric Effects.
Variables with "Pr(F) \> 0.05" suggest non-linear relationships.

From the output:

The variable "Private" does not appear in the "Nonparametric Effects"
section because it is a binary categorical variable.

There is strong evidence of non-linear relationships between "Outstate"
and "Expend" (very strong), "Accept", "Grad.Rate", "F.Undergrad"

There is marginal evidence for "Room.Board"

Other variables, including "Apps", "Top10perc", "perc.alumni", and
"Terminal", do not show significant non-linear effects — a linear fit is
sufficient.

While the smoothing plot for "Terminal" shows apparent curvature, the
corresponding nonparametric p-value (p = 0.206) suggests that this
nonlinearity is not statistically significant. This indicates that the
observed curve might be due to random variation or overfitting, rather
than reflecting a true nonlinear relationship in the population.

```{r 10 (d)}
summary(fit_gam)
```
