---
title: "hw7_XQ"
author: "XianjieQue"
date: "2025-04-08"
output: html_document
---

### Conceptual: 1 

I need to prove that the logistic function:

$$
p(X) = \frac{e^{\beta_0 + \beta_1 X}}{1 + e^{\beta_0 + \beta_1 X}} \tag{4.2}
$$

is equivalent to the logit representation:

$$
\frac{p(X)}{1 - p(X)} = e^{\beta_0 + \beta_1 X} \tag{4.3}
$$

Start from equation (4.2):

$$
p(X) = \frac{e^{\beta_0 + \beta_1 X}}{1 + e^{\beta_0 + \beta_1 X}}
$$

Compute (1 - p(X)):

$$
1 - p(X) = 1 - \frac{e^{\beta_0 + \beta_1 X}}{1 + e^{\beta_0 + \beta_1 X}} = 
\frac{1 + e^{\beta_0 + \beta_1 X} - e^{\beta_0 + \beta_1 X}}{1 + e^{\beta_0 + \beta_1 X}} =
\frac{1}{1 + e^{\beta_0 + \beta_1 X}}
$$

Take the ratio:

$$
\frac{p(X)}{1 - p(X)} =
\frac{\frac{e^{\beta_0 + \beta_1 X}}{1 + e^{\beta_0 + \beta_1 X}}}{\frac{1}{1 + e^{\beta_0 + \beta_1 X}}} =
e^{\beta_0 + \beta_1 X}
$$

Thus, 

$$
\frac{p(X)}{1 - p(X)} = e^{\beta_0 + \beta_1 X}
$$

is equivalent to the logistic function in equation (4.2).


## The attached question

### Split the data set into a training set and a test set.

```{r split}
library(glmnet)
library(pROC)

singh <- read.csv("singh2002.csv", header = TRUE)
dim(singh)

# check if there is missing data
sum(is.na(singh))

# head(singh, 10)

# convert y to numeric (0/1)

singh$y <- ifelse(singh$y == "cancer", 1, 0)

# split 2/3 as training, 1/3 as test
set.seed(17)
train <- sample(1:nrow(singh), 2/3*nrow(singh))

singh_train <- singh[train,]
singh_test <- singh[-train,]

x <- singh[,-6034]
y <- singh[,6034]
x_train <- as.matrix(singh_train[,-6034])
y_train <- as.numeric(singh_train$y)
x_test <- as.matrix(singh_test[,-6034])
y_test <- as.numeric(singh_test$y)
```

### (a) lasso

the classification error (AUC) is 0.8715

the number of non-zero coefficient estimates using the whole dataset (train + test) is 71

```{r (a) lasso}
set.seed(17)
lasso_model <- glmnet(x_train, y_train, family = "binomial", 
                      alpha = 1)
cv_out_a <- cv.glmnet(x_train, y_train, family = "binomial", 
                      alpha = 1)

## use labmda.min to get AUC

best_lam_1 <- cv_out_a$lambda.min

lasso_pred <- predict(lasso_model, s = best_lam_1, 
                      newx = x_test, type = "response")

auc_lasso <- roc(y_test, as.vector(lasso_pred))$auc
auc_lasso

## use the whole dataset (train + test) to get non_zero_coef

out_lasso <- glmnet(x, y, family = "binomial", alpha = 1)

lasso_coef_all <- predict(out_lasso, type = "coefficients", 
                      s = best_lam_1)[1:ncol(singh),]

non_zero_coef_all_1 <- lasso_coef_all[lasso_coef_all != 0]

print(paste("non_zero_coef_all", (length(non_zero_coef_all_1) - 1)))  # subtract 1 to exclude the intercept
```

### (b) elastic net

the classification error (AUC) is 0.9062

the number of non-zero coefficient estimates using the whole dataset (train + test) is 148

```{r (b) elastic net}
set.seed(17)
enet_cv <- cv.glmnet(x_train, y_train, family = "binomial", 
                     alpha = 0.5, nfolds = 10)
best_lambda_enet <- enet_cv$lambda.min
enet_model <- glmnet(x_train, y_train, family = "binomial", 
                     alpha = 0.5, lambda = best_lambda_enet)

# Make predictions and evaluate
enet_pred <- predict(enet_model, newx = x_test,
                     type = "response")

auc_enet <- roc(y_test, as.vector(enet_pred))$auc
auc_enet

## use the whole dataset (train + test) to get non_zero_coef

out_enet = glmnet(x, y, family = "binomial", alpha = 0.5)

enet_coef_all <- predict(out_enet, type = "coefficients", 
                      s = best_lambda_enet)[1:ncol(singh),]

non_zero_coef_all_2 <- enet_coef_all[enet_coef_all != 0]

print(paste("non_zero_coef_all", (length(non_zero_coef_all_2) - 1)))  # subtract 1 to exclude the intercept
```

### (c) SCAD

the classification error (AUC) is 0.8368

the number of non-zero coefficient estimates using the whole dataset (train + test) is 53

```{r (c) SCAD}
library(ncvreg)

# set.seed(400)
scad_cv <- cv.ncvreg(x_train, y_train, family = "binomial", 
                     penalty = "SCAD")
best_lambda_scad <- scad_cv$lambda.min
scad_model <- ncvreg(x_train, y_train, family = "binomial", 
                     penalty = "SCAD")

# Make predictions and evaluate
scad_pred <- predict(scad_model, x_test, type = "response",
                     lambda = best_lambda_scad)

auc_scad <- roc(y_test, as.vector(scad_pred))$auc
auc_scad

## use the whole dataset (train + test) to get non_zero_coef

out_scad <- ncvreg(x, y, family = "binomial", penalty = "SCAD")

scad_coef_all <- predict(out_scad, type = "coefficients",
                     lambda = best_lambda_scad)[1:ncol(singh),]

non_zero_coef_all_3 <- scad_coef_all[scad_coef_all != 0]

print(paste("non_zero_coef_all", (length(non_zero_coef_all_3) - 1)))  # subtract 1 to exclude the intercept
```

### (d) MCP

the classification error (AUC) is 0.8542

the number of non-zero coefficient estimates using the whole dataset (train + test) is 15

```{r (d) MCP}
set.seed(17)
mcp_cv <- cv.ncvreg(x_train, y_train, family = "binomial", 
                    penalty = "MCP")
best_lambda_mcp <- mcp_cv$lambda.min
mcp_model <- ncvreg(x_train, y_train, family = "binomial", 
                    penalty = "MCP")
mcp_model$lambda

# Make predictions and evaluate
mcp_pred <- predict(mcp_cv, x_test, type = "response")

auc_mcp <- roc(y_test, as.vector(mcp_pred))$auc
auc_mcp

## use the whole dataset (train + test) to get non_zero_coef

out_mcp <- ncvreg(x, y, family = "binomial", penalty = "MCP")

mcp_coef_all <- predict(out_mcp, type = "coefficients",
                        lambda = best_lambda_mcp)[1:ncol(singh),]

non_zero_coef_all_4 <- mcp_coef_all[mcp_coef_all != 0]

print(paste("non_zero_coef_all", (length(non_zero_coef_all_4) - 1)))  # subtract 1 to exclude the intercept
```

### (e) ISIS

the classification error (AUC) is 0.7951

the number of non-zero coefficient estimates using the whole dataset (train + test) is 5

```{r (e) ISIS}
library(SIS)
set.seed(17)

isis_model <- SIS(x_train, y_train, family = "binomial", 
                  tune = "cv", nfolds = 10, seed = 17)
best_lambda_isis <- isis_model$lambda

# Make predictions and evaluate
isis_pred <- predict(isis_model, x_test, type = "response",
                     lambda = best_lambda_isis)

auc_isis <- roc(y_test, as.vector(isis_pred))$auc
auc_isis

## use the whole dataset (train + test) to get non_zero_coef

out_isis = SIS(as.matrix(x), y, family = "binomial")

isis_coef_all <- predict(out_isis, type = "coefficients", 
                      s = best_lambda_isis)
non_zero_coef_all_5 <- isis_coef_all[isis_coef_all != 0]
print(paste("non_zero_coef_all", (length(non_zero_coef_all_5) - 1)))  # subtract 1 to exclude the intercept
```



