---
title: "DataMining_hw1_XianjieQue"
output: html_document
date: "2025-01-28"
---
### 1

### 1-(a): a flexible method is better.

When the sample size is large and the number of predictors is small, a flexible method can capture complex relationships in the data without overfitting.

### 1-(b): a inflexible method is better.

Flexible methods require a very large number of observations, which are not fit when the number of observations is small. They are prone to overfitting or randomly associating some predictors.

### 1-(c): a flexible method is better.

Because in this case, a inflexible method will lead to higher bias and may not be adequate. Flexible methods, such as decision trees or neural networks, can capture complex, non-linear relationships between predictors and the response.

### 1-(d): a inflexible method is better.

When the variance of the error terms is high, the data is noisy, and a flexible method may fit the noise in the data, leading to overfitting. Inflexible methods are more robust to noise.


### 2

### 2-(a): regression, inference, n = 500, p = 3

The CEO salary is a quantitative variable so it is a regression problem. The interest is to understand which factor affect CEO salary so we are more interested in inference.

### 2-(b): classification, prediction, n = 20, p = 13

We wish to know whether it will be a success or a failure. It is a qualitative response so can be referred as a classification problem. Since we want to know the result, the interest is prediction.

### 2-(c): regression, prediction, n = 52, p = 3

We want to predict the % change in the USD/Euro exchange. Obviously, we are more interested in prediction. Since the % change is a quantitative response, it is a regression problem.


### 5 

### Advantages of a flexible approach: 

1) less bias; 
2) can generate a much wider range to estimate f

### Disadvantages of a flexible approach: 

1) harder to interpret;
2) need a very large number of observations
3) more variance and may lead to overfit

When inference is the goal / there are a very large number of observations /we have non-linear patterns, a more flexible approach will be preferred.

When inference is the goal / there are few observations, a less flexible approach will be preferred.

### 6 The difference, advantages and disadvantages between parametric and non-parametric methods.

### Difference:
Parametric methods is based on model and parameters in the model. It simplifies the problem of estimating f down to estimating a set of parameters. Non-parametric methods don't assume explicit model or parameters, instead they seek an estimate of f that gets as close to the data points as possible without being too rough or wiggly.

### Advantage of Parametric methods (as opposed to non-parametric approaches):

1) Parametric methods don't need a very large number of observations.
2) It is more interpretable because the response dats is generated from model.

### Disadvantage of Parametric methods (as opposed to non-parametric approaches):

1) If the functional form used to estimate f is very different from the true f, the resulting model will not fit the data well.
2) It is less flexible.

### 8-(a)
```{r 8-(a)}
college <- read.csv("/Users/quexianjie/Desktop/1/SP2025.660 Biomedical Data Mining/hw1/College.csv", header = T, na.strings = "?", stringsAsFactors = T)
head(college)
dim(college)
```
### 8-(b)
```{r 8-(b)}
rownames(college) <- college[, 1]
View(college)

college <- college[, -1] 
View(college)
```
### 8-(c)-i
```{r 8-(c)-i}
summary(college)
```

### 8-(c)-ii

```{r 8-(c)-ii}
# a scatterplot matrix of the first ten columns
pairs(college[,1:10])
```

### 8-(c)-iii

```{r 8-(c)-iii}
attach(college)
plot(Private, Outstate,
     xlab = 'Private Indicator',
     ylab = 'Out-of-state tuition',
     main = 'Boxplots of Outstate versus Private'
     )
```

### 8-(c)-iv

```{r 8-(c)-iv}
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes" 
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)

summary(Elite)  # There are 78 elite universities

plot(college$Elite, college$Outstate,
     xlab = 'Elite Universities',
     ylab = 'Out-of-state tuition',
     main = 'Boxplots of Outstate versus Elite'
     )
```

### 8-(c)-v

```{r 8-(c)-v}
par(mfrow = c(2, 2))

hist(college$Outstate, 
     main = "Histogram_Outstate_10 bins", 
     xlab = "Outstate Tuition", 
     col = "skyblue", 
     breaks = 10)

hist(college$Room.Board, 
     main = "Histogram_Room and Board_20 bins", 
     xlab = "Room and Board Cost", 
     col = "lightgreen", 
     breaks = 20)

hist(college$Books, 
     main = "Histogram_Books_15 bins", 
     xlab = "Books Cost", 
     col = "lightpink", 
     breaks = 15)

hist(college$Personal, 
     main = "Histogram_Personal Expenses_25 bins", 
     xlab = "Personal Expenses", 
     col = "lightcoral", 
     breaks = 25)

par(mfrow = c(1, 1))
```
### 8-(c)-vi

A brief summary: 

From the correlation between variables, we know that "Number of full-time undergraduates" has a strong positive correlation with Enrolled number, Accepted number and Applications number, but has almost nothing to do with "Instructional expenditure per student". 

At the same time, chi-square analysis is performed on the remaining two qualitative variables. Since the P-value is less than 0.05, whether a college is Private and Elite is not random.

```{r 8-(c)-vi}
pairs(college)
cor(college[, sapply(college, is.numeric)])
chisq.test(college$Private, college$Elite)
```

### 9-(a)

Answer: From the result of "summary(Auto)", there are 9 predictors. "Name" is qualitative and the other 8 are quantitative. They are: "mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "year", "origin".

```{r 9-(a)}
# load the Auto data set and remove missing values
Auto <- read.table("/Users/quexianjie/Desktop/1/SP2025.660 Biomedical Data Mining/hw1/Auto.data", 
                   header = T, na.strings = "?", stringsAsFactors = T)
dim(Auto)
Auto <- na.omit(Auto) 
dim(Auto)

summary(Auto)
```

### 9-(b)

```{r 9-(b)}
# create a vector "quanti_x" including all 8 quantitative predictors
x <- colnames(Auto)
quanti_x <- setdiff(x, "name")

# use a for loop to calculate the range for each predictor
ranges <- list()
for (i in quanti_x){
  ranges[[i]] <- range(Auto[[i]])
}
print(ranges)
```

### 9-(c)

```{r 9-(c)}
# use a for loop to calculate the mean for each predictor
means <- list()
for (i in quanti_x){
  means[[i]] <- mean(Auto[[i]])
}
print(means)

# use a for loop to calculate the standard deviation for each predictor
sds <- list()
for (i in quanti_x){
  sds[[i]] <- sd(Auto[[i]])
}
print(sds)
```

### 9-(d)

```{r 9-(d)}
# remove the 10th through 85th observations
Auto_2 <- Auto[-c(10:85),]
dim(Auto_2)

# use the for loop to calculate the range again
ranges_2 <- list()
for (i in quanti_x){
  ranges_2[[i]] <- range(Auto_2[[i]])
}
print(ranges_2)

# use the for loop to calculate the mean again
means_2 <- list()
for (i in quanti_x){
  means_2[[i]] <- mean(Auto_2[[i]])
}
print(means_2)

# use the for loop to calculate the standard deviation again
sds_2 <- list()
for (i in quanti_x){
  sds_2[[i]] <- sd(Auto_2[[i]])
}
print(sds_2)
```

### 9-(e)

Comments: 
From the scatterplot matrix created by pairs() function, among these variables, "mpg" is strongly influenced by "displacement", "horsepower", and "weight", all showing a negative correlation with fuel efficiency.
Besides, "displacement" and "horsepower" have a strong positive correlation.

```{r 9-(e)}
attach(Auto)
cylinders <- as.factor(cylinders)
plot(cylinders, mpg, col = "red", varwidth = T,
    xlab = "cylinders", ylab = "MPG")

pairs(
 ~mpg + cylinders + displacement + horsepower + weight + acceleration
 + year + origin + name, data = Auto)
```

### 9-(f)

Answer: 
Yes. From the scatterplot matrix created by pairs() function, among these variables, "mpg" is strongly influenced by "displacement", "horsepower", and "weight", all showing a negative correlation with fuel efficiency. The relationship between "mpg" and other variables are more complex and possibly non-linear.
So we can use "displacement", "horsepower", and "weight" to predict "mpg". The greater "displacement", "horsepower", and "weight" are, the less "mpg" is.

```{r 9-(f)}
pairs(
 ~mpg + cylinders + displacement + horsepower + weight + acceleration
 + year + origin + name, data = Auto)

cor(Auto[, sapply(Auto, is.numeric)])
```
